\section{Variants and Results}\label{sec:results}

Some of the variations below include a Canny edge detector as a preprocessing
step. The parameter $\sigma$ determines the size of the gaussian kernel used by
the Canny algorithm. A value of $\sigma = 1.5$ was determined to be appropriate
for the image dataset used in the benchmark. Value larger than $\sigma = 2$
tend not to detect any edges, while smaller values produced more "false" edges
resulting from noise in the images.

The SEGMENT preprocessing step could unfortunately not use the same $1024
\times 768$ pixel images as inputs as the other preprocessors, because the
implementation provided by Arbelaez et al.\ \autocite{arbelaez_contour_2011}
could not be run on any machine available due to extreme memory requirements.
Therefore, the images in the database are rescaled to $320 \times 240$ pixels
for the evaluation of this algorithm. This might lead to loss of small features
    and puts limits on some parameters like the grid size $G$.

All pipeline variants utilize the FDCT to extract curve information. The number
of scales and the number of angles at the coarsest scale will be called $N_s$
and $N_{\theta}$ respectively. Based on experimentation and other publications
using the curvelet transform \autocite{mandal_curvelet_2009}
\autocite{guha_curvelet_????}, 4 scales and 12 angles are used in almost all
cases, since larger values did not consistently lead to better results. The
size of the codebook is set at 1000 visual words, as values beyond that did
not show improvements in evaluations performed by Nowak at al.\
\autocite{nowak_sampling_2006}. Eitz et al.\ \autocite{eitz_sketch-based_2010}
reported similar optimal values of 500 to 1000 visual words and attributed the
low number to the sparsity of edge-like features in sketches.

Throughout the correlation coefficient graphs in the following sections, the
red line denotes the best result obtained by Eitz et al.\ in
\autocite{eitz_sketch-based_2010} using the SHOG descriptor.

\input{chapters/chapter04/results_global.tex}
\input{chapters/chapter04/results_local.tex}


